NumPy---Array Programming in Python

### 2020-02-13 body.tex
Edit ecosystem section

Python is an open-source, general-purpose, interpreted programming language
well-suited to standard programming tasks such as cleaning data,
interacting with web resources, and parsing text.
Adding fast array operations and linear algebra allows scientists to do all
their work within a single language---and one that has the advantage of
being famously easy to learn and teach.
In fact, Python's adoption as a primary learning language in universities
has further popularized NumPy as a tool of choice for modern data science.

@@

SciPy and Matplotlib are tightly coupled with NumPy---in terms of
history, development, and use.
SciPy provides fundamental algorithms for scientific computing,
including mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The combination of NumPy, SciPy, and Matplotlib, together with
an advanced interactive environment like IPython \cite{perez2007ipython},
or Jupyter \cite{Kluyver:2016aa}, provides a solid foundation for array
programming in Python.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this foundation to provide several, widely used \emph{technique specific}
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
that in turn underlay numerous \emph{domain specific} projects
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,sunpy2015,2018EGUGA..2012146H}.
NumPy, as the focal point for the ecosystem of array-aware libraries,
sets documentation standards, and provides array testing infrastructure
and build support for integrating with Fortran code and using specialized
compilers and computing platforms.
% FIXME: what are "specialized compilers and computing
% platforms"? Does Fortran need to be mentioned here?

Many research groups have designed large,
complex scientific libraries, which add \emph{application specific} functionality
to the ecosystem.
For example, the \code{eht-imaging} library \cite{chael2019ehtim} developed by
the Event Horizon Telescope collaboration for radio interferometry imaging,
analysis, and simulation relies on many lower-level components of the scientific Python
ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image
reconstruction.
SciPy is used to provide tools used in general image processing such as
filtering and image alignment, while scikit-image, an image processing
library that extends SciPy, provides higher-level functionality such as
edge filters and Hough transforms.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} module for handling mathematical optimization tasks.
NetworkX \cite{SciPyProceedings_11}, a package for complex
network analysis, is used for the analysis of consistency among image
comparisons.
The community-developed astronomy library, Astropy \cite{astropy:2013, astropy:2018},
is used throughout \code{eht-imaging}, primarily to handle standard
astronomical file formats and time/coordinate transformations.
Matplotlib is used for visualizing data throughout the analysis pipeline,
including for the generation of the final image of the black hole.

The interactive environment created by this array programming
foundation---along with the surrounding ecosystem of tools---inside of
IPython or Jupyter, is ideally suited to exploratory data analysis.
Users fluidly inspect, manipulate, and visualize their data, and
rapidly iterate to refine programming statements. These statements are
then stitched together into imperative or functional programs, or
notebooks containing both computation and narrative.
Scientific computing beyond exploratory work is often done in a text editor
or an integrated development environment (IDEs) such as Spyder.
This rich and productive environment has made Python popular
for scientific research.

% this is the software engineering side of things
% as opposed to the interactive, exploration described above
NumPy is a community of practice with a culture of
employing software engineering practice to improve collaboration and
reduce error \cite{millman2014developing}.  This culture is not only
adopted by leaders in the project but also enthusiastically taught to
newcomers. The NumPy team was early in adopting distributed revision
control and code review to improve collaboration on code, and
continuous testing that runs an extensive battery of automated tests for
every proposed change to NumPy.  The project has comprehensive,
high-quality documentation, integrated with the source
code \cite{vanderwalt2008scipy,harrington2008scipy,harrington2009scipy}.

>>>>

Python is an open-source, general-purpose, interpreted programming language
well-suited to standard programming tasks such as cleaning data,
interacting with web resources, and parsing text.
Adding fast array operations and linear algebra allows scientists to do all
their work within a single language---and one that has the advantage of
being famously easy to learn and teach, as witnessed by its adoption
as a primary learning language in many universities.

@@

SciPy and Matplotlib are tightly coupled with NumPy---in terms of
history, development, and use.
SciPy provides fundamental algorithms for scientific computing,
including mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The combination of NumPy, SciPy, and Matplotlib, together with
an advanced interactive environment like IPython \cite{perez2007ipython},
or Jupyter \cite{Kluyver:2016aa}, provides a solid foundation for array
programming in Python.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this foundation to provide several, widely used \emph{technique specific}
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
that in turn underlay numerous \emph{domain specific} projects
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,sunpy2015,2018EGUGA..2012146H}.
NumPy, as the focal point for the ecosystem of array-aware libraries,
sets documentation standards, provides array testing infrastructure,
and adds build support for Fortran and other compilers.

Many research groups have designed large,
complex scientific libraries, which add \emph{application specific} functionality
to the ecosystem.
For example, the \code{eht-imaging} library \cite{chael2019ehtim} developed by
the Event Horizon Telescope collaboration for radio interferometry imaging,
analysis, and simulation, relies on many lower-level components of the scientific Python
ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image
reconstruction.
SciPy supplies tools for general image processing tasks such as
filtering and image alignment, while scikit-image, an image processing
library that extends SciPy, provides higher-level functionality such as
edge filters and Hough transforms.
The \code{scipy.optimize} module performs mathematical optimization.
NetworkX \cite{SciPyProceedings_11}, a package for complex
network analysis, is used to verify image comparison consistency.
Astropy \cite{astropy:2013, astropy:2018} handles standard
astronomical file formats and computes time/coordinate transformations.
Matplotlib is used to visualize data and to generate the final image of the black hole.

The interactive environment created by the array programming
foundation along with the surrounding ecosystem of tools---inside of
IPython or Jupyter---is ideally suited to exploratory data analysis.
Users fluidly inspect, manipulate, and visualize their data, and
rapidly iterate to refine programming statements. These statements are
then stitched together into imperative or functional programs, or
notebooks containing both computation and narrative.
Scientific computing beyond exploratory work is often done in a text editor
or an integrated development environment (IDEs) such as Spyder.
This rich and productive environment has made Python popular
for scientific research.

% this is the software engineering side of things
% as opposed to the interactive, exploration described above

To complement this facility for exploratory work and rapid
prototyping, NumPy has developed a culture of
employing time-tested software engineering practices to improve collaboration and
reduce error \cite{millman2014developing}.  This culture is not only
adopted by leaders in the project but also enthusiastically taught to
newcomers. The NumPy team was early in adopting distributed revision
control and code review to improve collaboration on code, and
continuous testing that runs an extensive battery of automated tests for
every proposed change to NumPy.  The project also has comprehensive,
high-quality documentation, integrated with the source
code \cite{vanderwalt2008scipy,harrington2008scipy,harrington2009scipy}.
